---
title: "BayCauRETM"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/LnnnnYW/BayCauRETM/actions/workflows/coverage.yaml/badge.svg)](https://github.com/LnnnnYW/BayCauRETM/actions/workflows/coverage.yaml)
[![Coverage Status](https://coveralls.io/repos/github/LnnnnYW/BayCauRETM/badge.svg)](https://coveralls.io/github/LnnnnYW/BayCauRETM)
![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)
<!-- badges: end -->





**BayCauRETM** (Bayesian Causal Recurrent-Event and Terminal-Event Modeling)  
provides a *fully Bayesian* workflow for **discrete-time causal inference**  
with recurrent-event counts jointly modeled with a terminal event.

- Canonical long-format preprocessing with automated lags  
- Stan-based joint modeling (`fit_causal_recur()`) with gAR(1) priors  
- Posterior *g*-computation for alternative treatment-start strategies  
- Diagnostics: MCMC convergence, propensity-score, switching-hazard  
- Publication-ready tables & plots

## Installation

``` r
# Install development version from GitHub
# install.packages("pak")
pak::pak("LnnnnYW/BayCauRETM")

# Or with devtools:
# devtools::install_github("LnnnnYW/BayCauRETM")
```

## Dependencies

The following R packages are required for `BayCauRETM`:

- [rstan](https://cran.r-project.org/package=rstan)
- [dplyr](https://cran.r-project.org/package=dplyr)
- [ggplot2](https://cran.r-project.org/package=ggplot2)
- [bayesplot](https://cran.r-project.org/package=bayesplot)
- [stats](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/00Index.html) *(base package)*
- [magrittr](https://cran.r-project.org/package=magrittr)
- [rlang](https://cran.r-project.org/package=rlang)
- [gt](https://cran.r-project.org/package=gt)
- [knitr](https://cran.r-project.org/package=knitr)
- [plotly](https://cran.r-project.org/package=plotly)
- [tidyr](https://cran.r-project.org/package=tidyr)
- [writexl](https://cran.r-project.org/package=writexl)

These are listed in the `Imports` field of the `DESCRIPTION` file and will be installed automatically with the package.


## Example

```{r full-workflow, eval = FALSE}
library(BayCauRETM)

## 1. Simulate toy data: 4 subjects × 3 intervals
set.seed(123)
toy <- data.frame(
  pat_id = rep(1:4, each = 3),
  k_idx  = rep(1:3, 4),
  Y_obs  = rpois(12, 1),
  T_obs  = rbinom(12, 1, 0.3),
  A      = rbinom(12, 1, 0.4)
)

## 2. Pre-process: add lags (Y_prev, T_prev) and fill missing intervals
pre <- preprocess_data(toy, K = 3)

## 3. Fit Bayesian joint model (tiny MCMC just for illustration)
fit <- fit_causal_recur(
  data       = pre$processed_df,
  K          = 3,
  formula_T  = T_obs ~ Y_prev + A + k_idx,
  formula_Y  = Y_obs ~ Y_prev + A + k_idx,
  prior      = list(
                 eta_beta  = 0, sigma_beta  = 1, rho_beta   = 0.5,
                 eta_gamma = 0, sigma_gamma = 1, rho_gamma  = 0.5),
  num_chains = 1,
  iter       = 200,      # increase for real analysis
  verbose    = FALSE
)

## 4. MCMC convergence diagnostics
diag <- mcmc_diagnosis(fit)
print(diag)
# plot(diag)

## 5. Posterior g-computation: start at s = 1, 2, 3 vs never
gout <- g_computation(fit, s_vec = 1:3, B = 30)
print(gout)
# plot(gout, ref_line = 0)

## 6. Propensity-score diagnostics
psd <- propensity_score_diagnostics(
         data       = fit$data_preprocessed,
         treat_col  = "A",
         covariates = c("Y_prev", "k_idx"))
# plot(psd, type = "histogram")

## 7. Switching-hazard diagnostics
sw <- switching_probability_summary(fit$data_preprocessed)
# plot(sw)

## 8. Merge posterior and g-computation summaries
tbl <- result_summary_table(fit, gout, s_vec = 1:3, format = "kable")
print(tbl)

```
## Documentation & Demos

- Full demo: `inst/demo/demo_full.R`
- Help pages for each function:  
  `?preprocess_data`, `?fit_causal_recur`, `?g_computation`, etc.


## Reporting Issues

If you encounter bugs or have feature suggestions, please  
open an [issue on GitHub](https://github.com/LnnnnYW/BayCauRETM/issues).  
Pull requests are always welcome!

## Citation

If you use **BayCauRETM**, please cite:

## Contact

The corresponding package author are Yuqin Wang (email: yuqin_wang@brown.edu) and Arman Oganisian (email: arman_oganisian@brown.edu).


> Tip: For reproducibility, knit this README from a `README.Rmd` source  
> using `devtools::build_readme()`.

